---
title: "Intratumoral T-cell and B-cell receptor architecture associates with distinct immune tumor microenvironment features and clinical outcomes of anti-PD-1/L1 immunotherapy"
subtitle: <center> TCGA data preparation</center>
author: "Aimilia Schina"
date: '`r paste("First created on 9 October 2019. Updated on ", format(Sys.Date(), "%d %B %Y"))`'
output:
  html_document:
    css: style.css
    code_folding: hide
    #smooth_scroll: yes
    fig_caption: yes
    #highlight: textmate
    #theme: cerulean
    #theme: simplex
    toc: yes
    toc_float:
      collapsed: true
      smooth_scroll: true
    toc_depth: 6
    number_sections: true
    #df_print: paged

---


```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = 'D:/BIOINF/PROJECTS/2_ANALYSES/18_GIT_PAPERS/TCR_BCR_antiPD1')
unlink("scripts/main/A_06_tcga_clinical_biomarker_metadata_cache", recursive = TRUE)
```



```{r,warning=F, message=F}
#####################
### Load packages ###
#####################
library(pacman)
pacman::p_load(DT, dplyr,readxl,TCGAbiolinks,splitstackshape,RTCGA,stringr,tibble,
               VennDiagram, randomcoloR, Amelia, mlbench, hrbrthemes, ggpubr, superheat, estimate,extrafont,TCGAbiolinks,ggvenn,
               data.table)
extrafont::loadfonts(device="win")
windowsFonts(sans="Palatino Linotype")
loadfonts(device="win")
loadfonts(device="postscript")

```


```{r,warning=F, message=F}
###########################
### Main analysis paths ###
###########################
# Main
scriptsPath <- paste0("scripts/")
scriptsFunctionsPath <- paste0(scriptsPath,"functions/")
projectDataPath <- paste0("data/")
# Input
tcgaInputData <- paste0(projectDataPath,"TCGA/")
otherInputData <- paste0(projectDataPath,"other/")
referenceInputData <- paste0(projectDataPath,"reference/")
# Output
projectOutDataPath <- paste0("output/data_files/")
tcgaIntermediateData <- paste0(projectOutDataPath,"TCGA/")
# Session/dependencies
sessionInfoPath <- paste0("session_info/")

######################################
## Create intermediate output paths ##
######################################
if (!dir.exists(paste0(tcgaIntermediateData,"purity/"))) {dir.create(paste0(tcgaIntermediateData,"purity/"))}


##################################
### LOAD SOURCE FUNCTIONS FILE ###
##################################
source(paste0(scriptsFunctionsPath,"tcgaReplicateFilter.R"), local=T)
source(paste0(scriptsFunctionsPath,"tcga_clinical_biomarker_metadata_functions.R"), local=T)


#####################
### File suffixes ###
#####################
Rdata.suffix <- ".RData"


########################
### Script variables ###
########################
run_clean <- TRUE
```




# Load data {.tabset .tabset-fade .tabset-pills}

Loading basic data objects and files for analysis.

## Reference data

```{r,warning=F, message=F}
geneid2symb <- read.table(file.path(paste0(referenceInputData, "genID2name.csv")), header = TRUE, sep = ",")
geneid2symb$gene_id <- as.character(geneid2symb$gene_id)
geneid2symb$gene_name <- as.character(geneid2symb$gene_name)
```

## Gene expression signatures

TIS-GEP, TuTack, TLS (Johnsson paper) signatures.

```{r,warning=F, message=F}
## Immune Gene signatures
gep_weights_df <- read.table(paste0(otherInputData,"Tcell_inflamed_gep_weights.txt"), header = TRUE, sep = "\t" )
normalization_genes_df <- read.table(paste0(otherInputData,"normalization_genes.txt"), header = TRUE, sep = "\t" )

# Load TuTack gene signature
TuTACK.DF <-loadRData(paste0(otherInputData,"TuTack_signature_weights",Rdata.suffix))
# Load tls signature
tls_df <- read.table(paste0(otherInputData,"tls_genes.txt"), header = TRUE, sep = "\t" )

```

## Non filtered nonduplicated RNA TCGA sample data, including added molecular subtypes

```{r,warning=F, message=F}
###############
## TCGA DATA ##
###############

### Load non filtered nonduplicated RNA TCGA sample data, including added molecular subtypes
biospecimen.tpnt.rna.merged.cl.f<-loadRData(paste0(tcgaIntermediateData,"biospecimen.merged.tpnt.RNA.noDup.sub.RData"))
# contains aliquot, analyte, patient, sample barcode

```

## Other TCGA data from publications

```{r,warning=F, message=F}
## TMB data
tmbDf <- read.table(paste0(tcgaInputData,"mutation-load_updated.txt"),sep="\t",header=T, stringsAsFactors=F,check.names=F)
colnames(tmbDf) <- gsub(" ", "_",colnames(tmbDf) )
colnames(tmbDf) <- gsub("-", "_",colnames(tmbDf) )
# contains patient, sample barcode


## ORR data
orrDf <- read_excel(paste0(otherInputData,"ORR_TMB_table.xlsx"),sheet = 2)  
orrDf <- as.data.frame(orrDf[,c(1:4,9:10)])
colnames(orrDf) <- c("TCGA_ca", "PD1L1_responders","PD1L1_totalTreated", "ORR", "project","Additional_info")
# Select only lines with complete info in first 5 columns
orrDf_final <- orrDf[complete.cases(orrDf[,1:5]),]
# Change name for COAD MSI &MSS
orrDf_final$project[7:8] <- c("COAD_MSI", "COAD_MSS")
# only info for cohort

## MiTCR data
mitcrDf <- read.table(paste0(tcgaInputData, "samples_tcr_mitcr.txt"),sep="\t",header=T, stringsAsFactors=F,check.names=F)
mitcrDf$chain <- apply(mitcrDf[,c("V_segments","V_alleles","J_segments","J_alleles")],1, function(x) paste(x,collapse = " "))

mitcrDf$chain <- ifelse(grepl("TRA",mitcrDf$chain),"alpha",ifelse(grepl("TRB",mitcrDf$chain),"beta",ifelse(grepl("TRG",mitcrDf$chain),"gamma",ifelse(grepl("TRD",mitcrDf$chain),"delta",NA))))
# contains aliquot barcode


mitcrClonesDf <- plyr::count(mitcrDf, c("Sample_ID", "chain"))
colnames(mitcrClonesDf)[3] <- "NumClones"

## Try with ImmPancer new mitcr file found
# Load excel file
mitcrDf2 <- read_excel(paste0(tcgaInputData,"1-s2.0-S1074761318301213-mmc2.xlsx"),na = "NA")
colnames(mitcrDf2) <- gsub(" ","_",colnames(mitcrDf2))
colnames(mitcrDf2) <- gsub("\\...","_",colnames(mitcrDf2))
# contains patient barcode


## Participant metadata
participDf <- read.table(paste0(tcgaInputData,"participant.tsv"),sep="\t",header=T,stringsAsFactors=F,check.names=F)
participDf$sample_id2 <- sub('.*\\/', '', participDf$sample_id)
participDf$sample_id2 <- sub('\\..*', '', participDf$sample_id2)
# contains aliquot barcode


## Danaher immune infiltration data
### Danaher does not include results for UCEC!!!!-CHECK CIBERSORT
danaherImmInfDf <-read.table(paste0(tcgaInputData, "tcga_danaher.csv"),sep=",",header=T, stringsAsFactors=F,check.names=F)
colnames(danaherImmInfDf)[1] <- "Sample_ID"
danaherImmInfDf$Sample_ID <- gsub("\\.","-",danaherImmInfDf$Sample_ID)
# contains aliquot  barcode

## Cibersort data

## Clinical Survival data
# Load clinical data for tcga samples
#loading Clinical data https://gdc.cancer.gov/about-data/publications/PanCan-Clinical-2018
clinDf <- read.table(paste0(tcgaInputData,"TCGA-CDR-SupplementalTableS1_sheet.txt"),sep="\t",header=T, stringsAsFactors=F,check.names=F,comment.char="@")
# contains a patient barcode

## TCGA Tumor Purity - EVENTUALLY I CALCULATE MYSELF
tumorPurity <- read_excel(paste0(tcgaInputData,"41467_2015_BFncomms9971_MOESM1236_ESM.xlsx"),na = "NA",skip=3)
# dim(tumorPurity)
# unique(tumorPurity$`Cancer type`) %>% length()
tumorPurity <- tumorPurity[,1:3]
```

# Estimate tumor purity

Run once since takes a lot of time. To run again, set run_clean to TRUE

```{r,warning=F, message=F,fig.show = 'hold', eval = TRUE, fig.width=15, fig.height=12}

if(isTRUE(run_clean)){
  #####################################
  #####################################
  # Load per tissue and run ESTIMATE
  tcga_projects <-TCGAbiolinks:::getGDCprojects()$project_id
  mycancertypes.off <- tcga_projects[grepl("^TCGA-",tcga_projects)] # total 33
  
  res<- sapply(mycancertypes.off, function(x) calculateEstimateScore(paste0(tcgaInputData, "htseq_vst/"),"_RNASeqCountsVST.RData",paste0(tcgaIntermediateData,"purity/"),paste0(x),geneid2symb=geneid2symb))
  # # Just using the counts
  
  estimateDFs<- sapply(mycancertypes.off, function(x) readEstimateFiles(paste0(tcgaIntermediateData,"purity/"),x), simplify = FALSE)
  # estimateDFs.filt <- estimateDFs[-2]
  # Remove empty dataframes
  estimateDFs.filt <- Filter(function(x) !is.null(x), estimateDFs)
  # Merge into one dataframe
  estimateDFs.merged <- bind_rows(estimateDFs.filt)
  
  ## Select primary solid tumor codes
  estimateDFs.merged.tp <- subset(estimateDFs.merged,substr(estimateDFs.merged$run_accession,14,15) == "01")
  
  tumorPurityEstimate <- estimateDFs.merged.tp
  ## CALCULATE TUMOR PURITY
  tumorPurityEstimate$TumorPurity <- cos(0.6049872018+0.0001467884*tumorPurityEstimate$ESTIMATEScore)
  # Fix sample id
  tumorPurityEstimate$bcr_patient_barcode <- substr(tumorPurityEstimate$run_accession,1,12)
  tumorPurityEstimate$bcr_patient_barcode <- gsub(".", "-", tumorPurityEstimate$bcr_patient_barcode, fixed = TRUE)
  tumorPurityEstimate$run_accession <- gsub(".", "-", tumorPurityEstimate$run_accession, fixed = TRUE)
  
  # Save data
  save(tumorPurityEstimate, file=paste0(tcgaIntermediateData,"TCGA_estimateDF.notFiltered",Rdata.suffix))
}else{
  if(isTRUE(file.exists(paste0(tcgaIntermediateData,"TCGA_estimateDF.notFiltered",Rdata.suffix)))){
    tumorPurityEstimate <- loadRData(paste0(tcgaIntermediateData,"TCGA_estimateDF.notFiltered",Rdata.suffix))
  }else{
    print("There is no file with estimated tumor purity, set run_clean to True to calculate tumor purity for TCGA samples.")
  }
}


```

# Filter data for non duplicated TCGA samples

Make sure all files including the TCGA cohorts are uniform, same names:

* COAD.MSS : COAD_MSS
* COAD.MSI_High: COAD_MSI
* COAD.MSI_Low: COAD_MSI_low
* BRCA.TripleNegative : BRCA_TN
* BRCA.Her2: BRCA_Her2
* BRCA.LuminalA : BRCA_LuminalA
* BRCA.LuminalB :BRCA_LuminalB
* BRCA.Normal: BRCA_Normal


```{r,warning=F, message=F}
# ## In biοspecimen DF
biospecimen.tpnt.rna.merged.cl.f <-tcga_cohortTypesCorrection(biospecimen.tpnt.rna.merged.cl.f,"type")

## TMB data--NOT NEEDED
tmbDf <-tcga_cohortTypesCorrection(tmbDf,"Cohort")

## ORR data
orrDf_final<-tcga_cohortTypesCorrection(orrDf_final,"project")

## MiTCR data--NOT NEEDED
mitcrDf2 <-tcga_cohortTypesCorrection(mitcrDf2,"TCGA_Study")

## Participant metadata--NOT NEEDED
participDf <-tcga_cohortTypesCorrection(participDf,"disease")

## Danaher immune infiltration data--NOT NEEDED
### Danaher does not include results for UCEC!!!!-CHECK CIBERSORT
danaherImmInfDf <-tcga_cohortTypesCorrection(danaherImmInfDf,"tcga.dataset")

## Cibersort data

## Clinical Survival data--NOT NEEDED
clinDf <-tcga_cohortTypesCorrection(clinDf,"type")

# Tumor purity
tumorPurity <- tcga_cohortTypesCorrection(tumorPurity,"Cancer type")

```

We check the Tissue Source site Codes for the different data we have above, to see if they have common codes.

```{r,warning=F, message=F,fig.show = 'hold'}
biospTSS <- levels(as.factor(substr(biospecimen.tpnt.rna.merged.cl.f$bcr_patient_barcode,6,7)))
tmbTSS <-levels(as.factor(substr(tmbDf$Patient_ID,6,7)))
mitcrImmTSS <-levels(as.factor(substr(mitcrDf2$TCGA_Participant_Barcode,6,7)))
mitcrClonesTSS <-levels(as.factor(substr(mitcrClonesDf$Sample_ID,6,7)))
danaherTSS <-levels(as.factor(substr(danaherImmInfDf$Sample_ID,6,7)))
survTSS <-levels(as.factor(substr(clinDf$bcr_patient_barcode,6,7)))
participTSS <-levels(as.factor(substr(participDf$`entity:participant_id`,6,7)))
tumorPurityTSS <- levels(as.factor(substr(tumorPurityEstimate$run_accession,6,7)))
```

```{r,warning=F, message=F,error=F,echo=F,fig.show = 'hold',results = FALSE}
# colors
set.seed(1234)
palette <- distinctColorPalette(5)

vinput1 <- list(Biospecimen=biospTSS, 
               TMB=tmbTSS, 
               # MiTCRimm=mitcrImmTSS, 
               MiTCRclones=mitcrClonesTSS, 
               Danaher=danaherTSS, 
               Survival=survTSS#,
               #Purity = tumorPurityTSS
               )


vp <-invisible(venn.diagram(vinput1, 
                   alpha = 0.6, filename = NULL,lty = "blank", fill = palette,
                   main= "",
                   main.cex = 1.1,
                   main.fontfamily="sans",
                   fontfamily ="sans",
                   sub.fontfamily = "sans",
                   cat.fontfamily = "sans",disable.logging = TRUE, output=FALSE

));

grid.newpage()
grid.draw(vp)
```

```{r,warning=F, message=F,error=F,echo=F,fig.show = 'hold',results = FALSE}
# colors
set.seed(1234)
palette <- distinctColorPalette(5)

vinput1 <- list(Biospecimen=biospTSS, 
               TMB=tmbTSS, 
               # MiTCRimm=mitcrImmTSS, 
               MiTCRclones=mitcrClonesTSS, 
               Danaher=danaherTSS, 
               # Survival=survTSS#,
               Purity = tumorPurityTSS
               )


vp <- invisible(venn.diagram(vinput1, 
                   alpha = 0.6, filename = NULL,lty = "blank", fill = palette,
                   main= "",
                   main.cex = 1.1,
                   main.fontfamily="sans",
                   fontfamily ="sans",
                   sub.fontfamily = "sans",
                   cat.fontfamily = "sans",disable.logging = TRUE, output=FALSE

));

grid.newpage()
grid.draw(vp)
```


```{r,warning=F, message=F,error=F,echo=F,fig.show = 'hold',results = FALSE}
# colors
set.seed(1234)
palette <- distinctColorPalette(5)

vinput2 <- list(Biospecimen=biospTSS, 
               TMB=tmbTSS, 
               MiTCRimm=mitcrImmTSS, 
               # MiTCRclones=mitcrClonesTSS, 
               Danaher=danaherTSS, 
               Survival=survTSS)

vp <- invisible(venn.diagram(vinput2, 
                   alpha = 0.6, filename = NULL,lty = "blank", fill = palette,
                   main= "",
                   main.cex = 1.1,
                   main.fontfamily="sans",
                   fontfamily ="sans",
                   sub.fontfamily = "sans",
                   cat.fontfamily = "sans",disable.logging = TRUE, output=FALSE

));

grid.newpage()
grid.draw(vp)
```

```{r,warning=F, message=F,error=F,echo=F,fig.show = 'hold',results = FALSE}
# colors
set.seed(1234)
palette <- distinctColorPalette(5)

vinput3 <- list(Biospecimen=biospTSS, 
               TMB=tmbTSS, 
               MiTCRimm=mitcrImmTSS, 
               MiTCRclones=mitcrClonesTSS, 
               # Danaher=danaherTSS, 
               Survival=survTSS)

vp <- invisible(venn.diagram(vinput3, 
                   alpha = 0.6, filename = NULL,lty = "blank", fill = palette,
                   main= "",
                   main.cex = 1.1,
                   main.fontfamily="sans",
                   fontfamily ="sans",
                   sub.fontfamily = "sans",
                   cat.fontfamily = "sans",disable.logging = TRUE, output=FALSE

));

grid.newpage()
grid.draw(vp)
```

```{r,warning=F, message=F,fig.show = 'hold'}
# colors
set.seed(1234)
palette <- distinctColorPalette(2)

vinput4 <- list(#Biospecimen=biospTSS, 
               #TMB=tmbTSS, 
               MiTCRimm=mitcrImmTSS, 
               MiTCRclones=mitcrClonesTSS 
               # Danaher=danaherTSS, 
               #Survival=survTSS
               )

vp <- invisible(venn.diagram(vinput4, 
                   alpha = 0.6, filename = NULL,lty = "blank", fill = palette,
                   main= "",
                   main.cex = 1.1,
                   main.fontfamily="sans",
                   fontfamily ="sans",
                   sub.fontfamily = "sans",
                   cat.fontfamily = "sans",disable.logging = TRUE, output=FALSE

));

grid.newpage()
grid.draw(vp)
```

Loading TCGA project names
```{r,warning=F, message=F,fig.show = 'hold'}

# TCGA projects
# Find TCGA projects
tcga_projects <-TCGAbiolinks:::getGDCprojects()$project_id
mycancertypes.tcgaBiolinks <- tcga_projects[grepl("^TCGA-",tcga_projects)] # total 33
mycancertypes.tcgaBiolinks.ord <- mycancertypes.tcgaBiolinks[order(mycancertypes.tcgaBiolinks)]
#list of tumor types from biospecimen data:
mycancertypes.tcgaBiolinks.ord.clean <- gsub("TCGA-","",mycancertypes.tcgaBiolinks.ord)

#####################
## SELECT PROJECTS ##
#####################
# ## Exclude THYM and DLBC


```

Important to note that we have removed the THYM and DLBC tcga datasets. A rationale explanation of that is that these are "immune cell tumors" that do not fit our classification, as explained by Danaher et al. "the 3 cancers in TCGA arising from "immune" cells, i.e., thymoma, acute myeloid leukemia (AML), and diffuse large B-cell lymphoma (DLBC), all display expression patterns consistent with an effect of tumor type on the algorithm".

In addition we have added BRCA Triple Negative as a distinct subset of BRCA and we split COAD to COAD MSI (MSI high) and COAD MSS.

# Load biospecimen, HTSEQ-VST normalized and RSEM TCGA data

Load TCGA sample data
```{r,warning=F, message=F,fig.show = 'hold'}
# Assign tumor samples and normal samples
tcga_TP <- biospecimen.tpnt.rna.merged.cl.f[biospecimen.tpnt.rna.merged.cl.f$sample_type=="TP",] #selecting tumors
tcga_NT <- biospecimen.tpnt.rna.merged.cl.f[biospecimen.tpnt.rna.merged.cl.f$sample_type=="NT",] #selecting normal tissue

#list of tumor types from biospecimen data:
mycancertypes.biosp.molSubtypes <- unique(tcga_TP$type)
mycancertypes.biosp.molSubtypes.ord <- mycancertypes.biosp.molSubtypes[order(mycancertypes.biosp.molSubtypes)]

#####################
## SELECT PROJECTS ##
#####################
# ## Exclude THYM and DLBC
# Setting the cancertypes for orr data as well
mycancertypes.orr <- mycancertypes.biosp.molSubtypes.ord

# Let's summarize the TCGA biospecimen data we just loaded
# Number of patients per project
biospecimen.tpnt.rna.merged.patient.sub.table <-biospecimen.tpnt.rna.merged.cl.f %>% select(type,bcr_patient_barcode) %>% group_by(type) %>% distinct() %>% summarize(n())
# Number of aliquots per project
biospecimen.tpnt.rna.merged.aliq.sub.table <- biospecimen.tpnt.rna.merged.cl.f %>% select(type,bcr_aliquot_barcode) %>% group_by(type) %>% distinct() %>% summarise(n())

# Table of number of replicates per patients per project
biospecimen.tpnt.rna.merged.aliqPatient.sub.table <-biospecimen.tpnt.rna.merged.cl.f %>% select(type,bcr_patient_barcode, bcr_aliquot_barcode) %>% group_by(type,bcr_patient_barcode) %>% distinct() %>% summarize(n())
```


**Summary Tables of TCGA biospecimen data**

***

```{r, eval = TRUE}

datatable(biospecimen.tpnt.rna.merged.patient.sub.table, extensions = 'Buttons', options = list(
    dom = 'Bfrtip',
    buttons = c('copy', 'excel', 'csv' ),
    scrollX=TRUE,
    pageLength=15
  ),
  caption = 'Number of patients per project (legacy & harmonized)'
)
```

```{r, eval = TRUE}

datatable(biospecimen.tpnt.rna.merged.aliq.sub.table, extensions = 'Buttons', options = list(
    dom = 'Bfrtip',
    buttons = c('copy', 'excel', 'csv' ),
    scrollX=TRUE,
    pageLength=15
  ),
  caption = 'Number of aliquots per project (legacy & harmonized)'
)
```

```{r, eval = TRUE}

datatable(biospecimen.tpnt.rna.merged.aliqPatient.sub.table, extensions = 'Buttons', options = list(
    dom = 'Bfrtip',
    buttons = c('copy', 'excel', 'csv' ),
    scrollX=TRUE,
    pageLength=15
  ),
  caption = 'Number of replicate aliquots per patient per project (legacy & harmonized)'
)
```

## Calulcate TIS-GEP and TuTack signature scores

From the last table we can still see that there are several replicates for many patients. Eventually after merging with the expression data, we will filter the TCGA barcodes and end up with one aliquot per patient.

Load TCGA RNA HTSEQ-FPKM-UQ data for all cancers and calculate the signature scores for Expanded immune signature, IFNG (small, 6 genes),  T cell-inflamed GEP (18 genes),the HTSEQ-VST data to calculate our "TuTACK" gene signature and the VST normalized PD-L1 expression.

```{r,warning=F, message=F, echo=FALSE, results='hide',fig.show = 'hold'}
rownames(TuTACK.DF)<- str_replace(rownames(TuTACK.DF),"_","-")
TuTACK.DF$genes <- str_replace(TuTACK.DF$genes,"_","-")
TuTACK <- TuTACK.DF$genes
TuTACK_ensemble <- TuTACK.DF$ensemble
###############################################################
gene_signatures_rna <- data.frame(Sample_ID=character(),
tcga.dataset = character(),
IFNg_expanded_sigscore = numeric(),
IFNg_small_sigscore = numeric(),
TuTACK_sigscore = numeric(),
PD_L1 = numeric(),
TLS_score = numeric(),
stringsAsFactors=FALSE)

htseq_files <-list.files(path = paste0(tcgaInputData, "htseq_fpkm_uq/"), pattern = "_RNASeqFPKM_UQ.RData", all.files = FALSE,
full.names = FALSE, recursive = FALSE,
ignore.case = FALSE, include.dirs = FALSE, no.. = FALSE)
mycancertypes.htseq <- sub("_RNASeqFPKM_UQ.RData","",htseq_files)


# Get Ayers & TuTAck signature scores
# gene_signatures_rna <- sapply(mycancertypes.htseq,function(x) load_tcga_rna_add_signatureScore(selectedcancer=x,gene_signatures_rna, TuTACK_ensemble,gene_data_kal,paste0(tcgaInputData, "htseq_fpkm_uq/"),paste0(tcgaInputData, "htseq_vst/")),simplify = FALSE)

gene_signatures_rna <- sapply(mycancertypes.htseq,function(x) load_tcga_rna_add_signatureScore(selectedcancer=x,gene_signatures_rna, TuTACK_ensemble,tls_df,paste0(tcgaInputData, "htseq_fpkm_uq/"),paste0(tcgaInputData, "htseq_vst/")),simplify = FALSE)
  
###################################################
# Merge data
gene_signatures_rna  <-bind_rows(gene_signatures_rna , .id = "column_label")
colnames(gene_signatures_rna)[3] <- "tcga.dataset"

print("Number of aliquots in HTSEQ data")
dim(gene_signatures_rna)#9760
# Number of  unique patients
print("Number of unique patients in HTSEQ data")
length(unique(substr(gene_signatures_rna$Sample_ID,1,12)))#9679

```

Load the Level 3 RSEM-normalized RNASeqV2 data from TCGA

```{r,warning=F, message=F,fig.show = 'hold'}

# reading data
list.files( paste0(tcgaInputData, "rsem/")) %>% 
  file.path( paste0(tcgaInputData, "rsem/"), .) -> folder

folder %>%
  list.files %>%
  file.path( folder, .) %>%
  grep( pattern = 'illuminahiseq', x = ., value = TRUE) -> pathRNA

dataNames<- gsub(paste0(paste0(tcgaInputData, "rsem//"),"gdac.broadinstitute.org_"),"",pathRNA)
# str_remove(pathRNA,paste0(paste0(tcgaIntermediateData, "rsem//"),"gdac.broadinstitute.org_"))

dataNames<- unlist(lapply(strsplit(dataNames,".Merge_rnaseqv2"), `[[`, 1))
names(pathRNA) <- dataNames
```

Get datasets for all TCGA and calculate TIS score

```{r,warning=F, message=F}
if(isTRUE(run_clean)){
    gene_TISsignature_rna <- data.frame(Sample_ID=character(),
                                    tcga.dataset = character(),
                                    TIS_sigscore = numeric(),
                                    stringsAsFactors=FALSE)



  gene_TISsignature_rna <- sapply(mycancertypes.tcgaBiolinks.ord.clean,function(x) get_TIS_score_tcga_rsem(selectedcancer=x,gene_TISsignature_rna,gep_weights_df,normalization_genes_df),simplify = FALSE)
  
  
  # sapply(gene_TISsignature_rna, function(x) median(x[["TIS_sigscore"]]))
  # Merge data
  gene_TISsignature_rna  <-bind_rows(gene_TISsignature_rna , .id = "column_label")
  
  
  median(gene_TISsignature_rna$TIS_sigscore)
  
  # Save and don't run above each time
  save(gene_TISsignature_rna,file=paste0(tcgaIntermediateData,"gene_TISsignature_rna", Rdata.suffix))
  ##IMPORTANT!!: Select only primary solid tumor samples
  gene_TISsignature_rna_TP <- subset(gene_TISsignature_rna,substr(gene_TISsignature_rna$Sample_ID,14,15) == "01")
  # Histogram of TIS scores
  hist(gene_TISsignature_rna_TP$TIS_sigscore)
  ## CHECK
  print("Number of aliquots in RSEM Primary Tumor data")
  dim(gene_TISsignature_rna_TP)#9101
  # Number of  unique patients
  print("Number of unique patients in RSEM Primary Tumor data")
  length(unique(substr(gene_TISsignature_rna_TP$Sample_ID,1,12)))#9101
  
}else{
  if(isTRUE(file.exists(paste0(tcgaIntermediateData,"gene_TISsignature_rna", Rdata.suffix)))){
    gene_TISsignature_rna  <-loadRData(paste0(tcgaIntermediateData,"gene_TISsignature_rna", Rdata.suffix))
    ##IMPORTANT!!: Select only primary solid tumor samples
  gene_TISsignature_rna_TP <- subset(gene_TISsignature_rna,substr(gene_TISsignature_rna$Sample_ID,14,15) == "01")
  # Histogram of TIS scores
  hist(gene_TISsignature_rna_TP$TIS_sigscore)
  ## CHECK
  print("Number of aliquots in RSEM Primary Tumor data")
  dim(gene_TISsignature_rna_TP)#9101
  # Number of  unique patients
  print("Number of unique patients in RSEM Primary Tumor data")
  length(unique(substr(gene_TISsignature_rna_TP$Sample_ID,1,12)))#9101
    
  }else{
    print("There is no file with calculated TIS signatures, set run_clean to True to calculate TIS signature scores for TCGA samples.")
  }
}


```

The barcode missing (missing_barcodes) is a weird case, since when we look it up on [GDC portal](https://portal.gdc.cancer.gov/cases/9e18238f-ae85-4b75-ae87-d92da9731a40) it is annotated as a Recurrent Tumor, though in the barcode the sample type ID is 01 -which should be 02 for recurrent tumor.

We decide to exclude this barcode, since we do not know whether it is correct or not. Based on GDC portal it is a recurrent tumor, so it should be excluded. Also, by looking at the _merged_sample_quality_annotations.tsv_ file of the [PanCanAtlas initiative](https://gdc.cancer.gov/about-data/publications/pancanatlas) this aliquot is noted as: "Notification:Item in special subset:This is part of the ovarian triplet set." AND "Notification:Item is noncanonical:Ovarian Triplet. Sample TCGA-23-1023-01R is a recurrent sample that for historical reasons has retained the -01R designation rather than the conventional -02R designation used elsewhere in TCGA." It's not noted as a "do not use sample", but based on sample type id, it does not fit the requirements for the current analysis.


# MERGE ALL DATA

## Filtering TCGA barcodes & fixing data columns 

First rename column of aliquots so we can merge by one column, and transform data to have common columns with:

* project (BRCA, ACC)
* tcga.dataset (TCGA-BRCA, TCGA-ACC)
* bcr_aliquot_barcode#1:28
* bcr_analyte_barcode#1:20
* bcr_sample_barcode#1:16
* bcr_patient_barcode#1:12

We also need to add to all data apart from biospecimen the BRCA_Her2, BRCA_Luminal, BRCA_LuminalB,BRCA_Normal,BRCA_TN,COAD_MSI,COAD_MSI_low,COAD_MSS  (so duplicate samples will exist)

For some patients we have multiple samples, multiple vials etc. Multiple samples can exist because there are samples from primary solid tumors, solid tissue normal etc..

For tumor we only consider:

* Primary solid Tumor - 01
* (Recurrent Solid Tumor) -02
* (Additional - New Primary) - 05
* (Metastatic) -06
* (Additional Metastatic) -07

```{r,warning=F, message=F,echo=FALSE, results=FALSE,fig.show = 'hold', eval=TRUE}
filterTcgaBarcodes <- function(df,col){
  col.un <- as.name(col)
  df.barcodes <- tcga_replicateFilter(df[[col]] , analyte_target = "RNA",filter_FFPE=TRUE, full_barcode=TRUE)
  # length(tcga.tp.barcodes)
  # Subset to filtered barcodes
  df.final <- df %>% dplyr::filter(!!col.un %in% df.barcodes) %>% 
                                      distinct() %>% droplevels()
  df.final
}

# bcr_aliquot_barcode#1:28
# bcr_analyte_barcode#1:20
# bcr_sample_barcode#1:16
# bcr_patient_barcode#1:12
# project
# tcga.dataset
####

## BIOSPECIMEN DATA
biospecimen.tpnt.rna.merged.cl.f.sub <- biospecimen.tpnt.rna.merged.cl.f %>% dplyr::filter(sample_type=="TP")
colnames(biospecimen.tpnt.rna.merged.cl.f.sub)[1] <- c("tcga.dataset")
biospecimen.tpnt.rna.merged.cl.f.sub$project <-biospecimen.tpnt.rna.merged.cl.f.sub$type
biospecimen.tpnt.rna.merged.cl.f.sub$tcga.dataset<- paste0("TCGA-",biospecimen.tpnt.rna.merged.cl.f.sub$project)

biospecimen.tpnt.rna.merged.cl.f.sub <- biospecimen.tpnt.rna.merged.cl.f.sub %>% select(bcr_aliquot_barcode,
                                                              bcr_analyte_barcode,
                                                              bcr_sample_barcode,
                                                              bcr_patient_barcode,
                                                              project,
                                                              tcga.dataset,
                                                              sample_type_id,
                                                              sample_type,
                                                              analyte_type_id,
                                                              MSI_status) %>% distinct() %>% droplevels()


# Check duplicates
biospecimen.tpnt.rna.merged.cl.f.sub %>% dplyr:: filter(bcr_patient_barcode %in% biospecimen.tpnt.rna.merged.cl.f.sub$bcr_patient_barcode[which(duplicated(biospecimen.tpnt.rna.merged.cl.f.sub$bcr_patient_barcode))[1]])


biospecimen.tpnt.rna.merged.cl.f.sub.f <-filterTcgaBarcodes(biospecimen.tpnt.rna.merged.cl.f.sub,"bcr_aliquot_barcode")

# Check duplicate
biospecimen.tpnt.rna.merged.cl.f.sub.f %>% dplyr:: filter(bcr_patient_barcode %in% biospecimen.tpnt.rna.merged.cl.f.sub.f$bcr_patient_barcode[which(duplicated(biospecimen.tpnt.rna.merged.cl.f.sub.f$bcr_patient_barcode))[2]])
# We have patients with multiple aliquots, we should only care for patient and project
biospecimen.tpnt.rna.merged.cl.f.sub.f <- biospecimen.tpnt.rna.merged.cl.f.sub.f %>% select(bcr_patient_barcode, project, tcga.dataset) %>% distinct() %>% droplevels()

## HERE I HAVE MULTIPLE PATIENTS WITH MULTIPLE SAMPLES
## GENE SIGNATURES
gene_signatures_rna.sub <- gene_signatures_rna
colnames(gene_signatures_rna.sub)[1:2] <- c("project","bcr_aliquot_barcode")
gene_signatures_rna.sub$project <- str_remove(gene_signatures_rna.sub$project,"TCGA-")
gene_signatures_rna.sub$bcr_analyte_barcode <-substr(gene_signatures_rna.sub$bcr_aliquot_barcode,1,20)
gene_signatures_rna.sub$bcr_sample_barcode <-substr(gene_signatures_rna.sub$bcr_aliquot_barcode,1,16)
gene_signatures_rna.sub$bcr_patient_barcode <- substr(gene_signatures_rna.sub$bcr_aliquot_barcode,1,12)

# Select only TP
gene_signatures_rna.sub <-gene_signatures_rna.sub[which(substr(gene_signatures_rna.sub$bcr_aliquot_barcode,14,15)=="01"),]


gene_signatures_rna.sub <- gene_signatures_rna.sub %>% select(bcr_aliquot_barcode,
                                                              bcr_analyte_barcode,
                                                              bcr_sample_barcode,
                                                              bcr_patient_barcode,
                                                              project,
                                                              tcga.dataset,
                                                              IFNg_expanded_sigscore,
                                                              IFNg_small_sigscore,
                                                              TuTACK_sigscore,
                                                              PD_L1,
                                                              TLS_score) %>% distinct() %>% droplevels()

gene_signatures_rna.sub.f <-filterTcgaBarcodes(gene_signatures_rna.sub,"bcr_aliquot_barcode")

# Check duplicate
gene_signatures_rna.sub.f %>% dplyr:: filter(bcr_patient_barcode %in% gene_signatures_rna.sub.f$bcr_patient_barcode[which(duplicated(gene_signatures_rna.sub.f$bcr_patient_barcode))])

# YES PATIENT WITH TWO SAMPLES
# Since there mihgt be multiple vials for each patient, we aggregate by patient with the median of the values, BY CHAIN
####
# Aggregate vials/samples

gene_signatures_rna.sub.f.agg <- gene_signatures_rna.sub.f %>%
                      group_by( bcr_patient_barcode, project, tcga.dataset) %>%
                      summarise_at(c("IFNg_expanded_sigscore", "IFNg_small_sigscore", "TuTACK_sigscore" ,"PD_L1","TLS_score"),median, na.rm=TRUE) %>%
                      ungroup %>% distinct()
gene_signatures_rna.sub.f.agg %>% dplyr:: filter(bcr_patient_barcode %in% gene_signatures_rna.sub.f.agg$bcr_patient_barcode[which(duplicated(gene_signatures_rna.sub.f.agg$bcr_patient_barcode))])

# Check duplicate
gene_signatures_rna.sub.f.agg %>% dplyr:: filter(bcr_patient_barcode %in% gene_signatures_rna.sub.f.agg$bcr_patient_barcode[which(duplicated(gene_signatures_rna.sub.f.agg$bcr_patient_barcode))])

##########
gene_TISsignature_rna_TP.sub <- gene_TISsignature_rna_TP
colnames(gene_TISsignature_rna_TP.sub)[1:2] <- c("project","bcr_aliquot_barcode")
gene_TISsignature_rna_TP.sub$tcga.dataset <- paste0("TCGA-",gene_TISsignature_rna_TP.sub$tcga.dataset)
gene_TISsignature_rna_TP.sub$bcr_analyte_barcode <-substr(gene_TISsignature_rna_TP.sub$bcr_aliquot_barcode,1,20)
gene_TISsignature_rna_TP.sub$bcr_sample_barcode <-substr(gene_TISsignature_rna_TP.sub$bcr_aliquot_barcode,1,16)
gene_TISsignature_rna_TP.sub$bcr_patient_barcode <- substr(gene_TISsignature_rna_TP.sub$bcr_aliquot_barcode,1,12)

gene_TISsignature_rna_TP.sub <- gene_TISsignature_rna_TP.sub %>% select(bcr_aliquot_barcode,
                                                              bcr_analyte_barcode,
                                                              bcr_sample_barcode,
                                                              bcr_patient_barcode,
                                                              project,
                                                              tcga.dataset,
                                                              TIS_sigscore) %>% distinct() %>% droplevels()




gene_TISsignature_rna_TP.sub.f <-filterTcgaBarcodes(gene_TISsignature_rna_TP.sub,"bcr_aliquot_barcode")
# Check duplicate
gene_TISsignature_rna_TP.sub.f %>% dplyr:: filter(bcr_patient_barcode %in% gene_TISsignature_rna_TP.sub.f$bcr_patient_barcode[which(duplicated(gene_TISsignature_rna_TP.sub.f$bcr_patient_barcode))])


gene_TISsignature_rna_TP.sub.f <- gene_TISsignature_rna_TP.sub.f %>% select(-bcr_aliquot_barcode,
                                                              -bcr_analyte_barcode,
                                                              -bcr_sample_barcode) %>% distinct() %>% droplevels()
## TMB
tmbDf.sub <- tmbDf
colnames(tmbDf.sub)[1:3] <- c("project", "bcr_patient_barcode", "bcr_sample_barcode")
tmbDf.sub$tcga.dataset <- paste0("TCGA-",tmbDf.sub$project)

tmbDf.sub <- tmbDf.sub %>% select(bcr_sample_barcode,
                                  bcr_patient_barcode,
                                  project,
                                  tcga.dataset,
                                  Silent_per_Mb,
                                  Non_silent_per_Mb) %>% distinct() %>% droplevels()
tmbDf.sub$logTMB <- log(tmbDf.sub$Non_silent_per_Mb + 1)


# Select only TP
tmbDf.sub <-tmbDf.sub[which(substr(tmbDf.sub$bcr_sample_barcode,14,15)=="01"),]

# Check duplicate
tmbDf.sub %>% dplyr:: filter(bcr_patient_barcode %in% tmbDf.sub$bcr_patient_barcode[which(duplicated(tmbDf.sub$bcr_patient_barcode))])

tmbDf.sub <-tmbDf.sub %>% select(-bcr_sample_barcode) %>% distinct()

## TCR DATA
mitcrDf2.sub <- mitcrDf2
colnames(mitcrDf2.sub)[1:2] <- c("bcr_patient_barcode","project")
mitcrDf2.sub$tcga.dataset <- paste0("TCGA-",mitcrDf2.sub$project)
# It includes also survival data, need to see how to deal with that.
otherCols <- colnames(mitcrDf2.sub)[c(3,5:32,37:64)] #  EXCLUDE OS PFI
allCols <- c("bcr_patient_barcode","project",
             "tcga.dataset",
             "TCGA_Subtype",otherCols)
mitcrDf2.sub <- mitcrDf2.sub %>% select(all_of(allCols)) %>% distinct() %>% droplevels()

mitcrDf2.sub <- mitcrDf2.sub %>% select(-Eosinophils_61, -Neutrophils_60) %>% distinct() %>% droplevels()


# Check duplicate
mitcrDf2.sub %>% dplyr:: filter(bcr_patient_barcode %in% mitcrDf2.sub$bcr_patient_barcode[which(duplicated(mitcrDf2.sub$bcr_patient_barcode))])



#mitcrDf2.sub %>% dplyr::filter(complete.cases(OS)) %>% pull(OS) %>% length()

## Danaher
danaherImmInfDf.sub <-danaherImmInfDf

danaherImmInfDf.sub$project <- danaherImmInfDf.sub$tcga.dataset
danaherImmInfDf.sub$tcga.dataset <- paste0("TCGA-",danaherImmInfDf.sub$tcga.dataset)
colnames(danaherImmInfDf.sub)[1] <- c("bcr_aliquot_barcode")

danaherImmInfDf.sub$bcr_analyte_barcode <-substr(danaherImmInfDf.sub$bcr_aliquot_barcode,1,20)
danaherImmInfDf.sub$bcr_sample_barcode <-substr(danaherImmInfDf.sub$bcr_aliquot_barcode,1,16)
danaherImmInfDf.sub$bcr_patient_barcode <- substr(danaherImmInfDf.sub$bcr_aliquot_barcode,1,12)

otherCols <- colnames(danaherImmInfDf.sub)[c(3:16)]
allCols <- c("bcr_aliquot_barcode",
              "bcr_analyte_barcode",
              "bcr_sample_barcode",
              "bcr_patient_barcode",
              "project",
              "tcga.dataset",otherCols)
danaherImmInfDf.sub <- danaherImmInfDf.sub %>% select(all_of(allCols)) %>% distinct() %>% droplevels()


# Select only TP
danaherImmInfDf.sub <-danaherImmInfDf.sub[which(substr(danaherImmInfDf.sub$bcr_sample_barcode,14,15)=="01"),]

# Filter aliquots
danaherImmInfDf.sub.f <-filterTcgaBarcodes(danaherImmInfDf.sub,"bcr_aliquot_barcode")
# Check duplicate
danaherImmInfDf.sub.f %>% dplyr:: filter(bcr_patient_barcode %in% danaherImmInfDf.sub.f$bcr_patient_barcode[which(duplicated(danaherImmInfDf.sub.f$bcr_patient_barcode))])
# Aggregate vials/samples
danaherImmInfDf.sub.f.agg <- danaherImmInfDf.sub.f %>%
                      group_by( bcr_patient_barcode, project, tcga.dataset) %>%
                      summarise_at(colnames(danaherImmInfDf.sub.f)[7:ncol(danaherImmInfDf.sub.f)],median, na.rm=TRUE) %>%
                      ungroup %>% distinct()
danaherImmInfDf.sub.f.agg %>% dplyr:: filter(bcr_patient_barcode %in% danaherImmInfDf.sub.f.agg$bcr_patient_barcode[which(duplicated(danaherImmInfDf.sub.f.agg$bcr_patient_barcode))])




## Clincal DF
clinDf.sub <- clinDf
cols.num <- c("age_at_initial_pathologic_diagnosis",
                             "initial_pathologic_dx_year",
                              "birth_days_to",
                             "last_contact_days_to",
                             "death_days_to",
                             "new_tumor_event_dx_days_to",
                             "OS",
                             "OS.time",
                             "DSS", "DSS.time",
                             "PFI","PFI.time")
clinDf.sub[cols.num] <- sapply(clinDf.sub[cols.num],as.numeric)
# Remove first column
clinDf.sub <- clinDf.sub[,-1]
colnames(clinDf.sub)[2] <- "project" 

clinDf.sub$tcga.dataset <- paste0("TCGA-",clinDf.sub$project)

otherCols <- colnames(clinDf.sub)[c(3:33)]
allCols <- c("bcr_patient_barcode",
              "project",
              "tcga.dataset",otherCols)
clinDf.sub <- clinDf.sub %>% select(all_of(allCols)) %>% distinct() %>% droplevels()

# ,"age_at_initial_pathologic_diagnosis",
#                     "gender","race","ajcc_pathologic_tumor_stage",
#                     "birth_days_to", "vital_status", "tumor_status",
#                     "OS", "OS.time", "DSS", "DSS.time", "DFI", "DFI.time", "PFI", "PFI.time"

# Check duplicate
clinDf.sub %>% dplyr:: filter(bcr_patient_barcode %in% clinDf.sub$bcr_patient_barcode[which(duplicated(clinDf.sub$bcr_patient_barcode))])



## Participant DF
participDf.sub <- participDf
colnames(participDf.sub)[1:2] <- c("bcr_aliquot_barcode","project")

participDf.sub$tcga.dataset <- paste0("TCGA-",participDf.sub$project)
participDf.sub$bcr_analyte_barcode <-substr(participDf.sub$bcr_aliquot_barcode,1,20)
participDf.sub$bcr_sample_barcode <-substr(participDf.sub$bcr_aliquot_barcode,1,16)
participDf.sub$bcr_patient_barcode <- substr(participDf.sub$bcr_aliquot_barcode,1,12)


participDf.sub <- participDf.sub %>% select(bcr_aliquot_barcode,
                                            bcr_analyte_barcode,
                                            bcr_sample_barcode,
                                            bcr_patient_barcode,
                                            project,
                                            tcga.dataset,
                                            sample_type,
                                            disease_name) %>% distinct() %>% droplevels()

# Only TP
participDf.sub <-participDf.sub[which(substr(participDf.sub$bcr_sample_barcode,14,15)=="01"),]

# Filter TCGA aliquots
participDf.sub.f <-filterTcgaBarcodes(participDf.sub,"bcr_aliquot_barcode")
# Check duplicate
participDf.sub.f %>% dplyr:: filter(bcr_patient_barcode %in% participDf.sub.f$bcr_patient_barcode[which(duplicated(participDf.sub.f$bcr_patient_barcode))])

# We care about disease name
# Aggregate vials/samples
participDf.sub.f <- participDf.sub.f %>% select(bcr_patient_barcode, project, tcga.dataset, -sample_type, disease_name) %>%
                    distinct() %>% droplevels()


# Check duplicate
participDf.sub.f %>% dplyr:: filter(bcr_patient_barcode %in% participDf.sub.f$bcr_patient_barcode[which(duplicated(participDf.sub.f$bcr_patient_barcode))])


## TUMOR PURITY
tumorPurityEstimate.sub <- tumorPurityEstimate
tumorPurityEstimate.sub$TumorPurity <- cos(0.6049872018+0.0001467884*tumorPurityEstimate.sub$ESTIMATEScore)
colnames(tumorPurityEstimate.sub)[1] <-"bcr_aliquot_barcode"

# tumorPurityEstimate.sub$
# participDf.sub$tcga.dataset <- paste0("TCGA-",participDf.sub$project)
tumorPurityEstimate.sub$bcr_analyte_barcode <-substr(tumorPurityEstimate.sub$bcr_aliquot_barcode,1,20)
tumorPurityEstimate.sub$bcr_sample_barcode <-substr(tumorPurityEstimate.sub$bcr_aliquot_barcode,1,16)
tumorPurityEstimate.sub$bcr_patient_barcode <- substr(tumorPurityEstimate.sub$bcr_aliquot_barcode,1,12)
# str_before_nth(tumorPurityEstimate.sub$bcr_aliquot_barcode,"-",2)

tumorPurityEstimate.sub <- tumorPurityEstimate.sub %>% select(bcr_aliquot_barcode,
                                            bcr_analyte_barcode,
                                            bcr_sample_barcode,
                                            bcr_patient_barcode,
                                            StromalScore,
                                            ImmuneScore,
                                            ESTIMATEScore,
                                            TumorPurity) %>% distinct() %>% droplevels()

# Only TP
tumorPurityEstimate.sub <-tumorPurityEstimate.sub[which(substr(tumorPurityEstimate.sub$bcr_sample_barcode,14,15)=="01"),]

#Filter tCGA aliquots
tumorPurityEstimate.sub.f <-filterTcgaBarcodes(tumorPurityEstimate.sub,"bcr_aliquot_barcode")
# Check duplicate
tumorPurityEstimate.sub.f %>% dplyr:: filter(bcr_patient_barcode %in% tumorPurityEstimate.sub.f$bcr_patient_barcode[which(duplicated(tumorPurityEstimate.sub.f$bcr_patient_barcode))])
# Aggregate vials/samples
tumorPurityEstimate.sub.f.agg <- tumorPurityEstimate.sub.f %>%
                      group_by( bcr_patient_barcode) %>%
                      summarise_at(colnames(tumorPurityEstimate.sub.f)[5:ncol(tumorPurityEstimate.sub.f)],median, na.rm=TRUE) %>%
                      ungroup %>% distinct()
tumorPurityEstimate.sub.f.agg %>% dplyr:: filter(bcr_patient_barcode %in% tumorPurityEstimate.sub.f.agg$bcr_patient_barcode[which(duplicated(tumorPurityEstimate.sub.f.agg$bcr_patient_barcode))])



# Check duplicate
tumorPurityEstimate.sub.f.agg %>% dplyr:: filter(bcr_patient_barcode %in% tumorPurityEstimate.sub.f.agg$bcr_patient_barcode[which(duplicated(tumorPurityEstimate.sub.f.agg$bcr_patient_barcode))])

## ORR
orrDf_final.sub <- orrDf_final
orrDf_final.sub$tcga.dataset <- paste0("TCGA-",orrDf_final.sub$project)
orrDf_final.sub <- orrDf_final.sub %>% select(project,
                                              tcga.dataset,
                                              TCGA_ca,
                                              PD1L1_responders,
                                              PD1L1_totalTreated) %>% distinct() %>% droplevels()

# ### OS DATA - SELECT FROM WHERE
# mitcrDf2.sub %>% dplyr::filter(complete.cases(OS)) %>% pull(OS) %>% length()
# which(complete.cases(clinDf.sub$OS)) %>% length()
patientProject.DF <-rbind(biospecimen.tpnt.rna.merged.cl.f.sub %>% select(bcr_patient_barcode, project, tcga.dataset),
      clinDf.sub %>% select(bcr_patient_barcode, project, tcga.dataset),
      participDf.sub %>% select(bcr_patient_barcode, project, tcga.dataset),
      mitcrDf2.sub %>% select(bcr_patient_barcode, project, tcga.dataset),
      danaherImmInfDf.sub %>% select(bcr_patient_barcode, project, tcga.dataset),
      gene_TISsignature_rna_TP.sub %>% select(bcr_patient_barcode, project, tcga.dataset),
      gene_signatures_rna.sub %>% select(bcr_patient_barcode, project, tcga.dataset),
      tmbDf.sub %>% select(bcr_patient_barcode, project, tcga.dataset)#,
      # tumorPurityEstimate.sub %>% select(bcr_patient_barcode, project, tcga.dataset)
      ) %>% distinct()
dim(patientProject.DF)

# Try toassign project tcga.dataset
setdiff( tumorPurityEstimate.sub.f.agg$bcr_patient_barcode,patientProject.DF$bcr_patient_barcode)

`%notin%` <- Negate(`%in%`)
unique(patientProject.DF$project)
tumorPurityEstimate.sub.f.agg <- merge(tumorPurityEstimate.sub.f.agg,patientProject.DF %>% dplyr::filter(project %notin% c("BRCA_LuminalA", "BRCA_Her2", "BRCA_LuminalB", "BRCA_Normal", "BRCA_TN", "COAD_MSS", "COAD_MSI_low", "COAD_MSI")), by="bcr_patient_barcode",all.x=TRUE)
tumorPurityEstimate.sub.f.agg <- tumorPurityEstimate.sub.f.agg[,c(1,6,7,2:5)]
tumorPurityEstimate.sub.f.agg %>% dplyr:: filter(bcr_patient_barcode %in% tumorPurityEstimate.sub.f.agg$bcr_patient_barcode[which(duplicated(tumorPurityEstimate.sub.f.agg$bcr_patient_barcode))])
```


## Cibersort subsets aggregation for B cells, CD8 T cells, CD4 T cells.

```{r,warning=F, message=F,fig.show = 'hold', eval=TRUE, fig.width=15, fig.height=12}

compactPopulations <- c("B_Cells_Naive","B_Cells_Memory","Plasma_Cells","T_Cells_CD8","T_Cells_CD4_Naive"           ,"T_Cells_CD4_Memory_Resting","T_Cells_CD4_Memory_Activated", "T_Cells_Regulatory_Tregs",
                        "NK_Cells_Resting","NK_Cells_Activated",
                        # "Dendritic_Cells_Resting","Dendritic_Cells_Activated",
                        # "Monocytes",
                        "Eosinophils_41","Mast_Cells_Resting","Mast_Cells_Activated","Neutrophils_48"
                        # "Macrophages_M0", "Macrophages_M1","Macrophages_M2"
                        )
# bcells <-c("B.Cells.Memory","B.Cells.Naive","Plasma.Cells")
# cd8 <-c("T.Cells.CD8.Naive","T.Cells.CD8.Actived","T.Cells.CD8.Memory")
# cd4 <-c("T.Cells.CD4.Naive","T.Cells.CD4.Memory")
# treg <-"Treg.Cells"
# nkcell <-c("NK.Resting", "NK.Actived")
# dc <-c("DC.Actived", "DC.Immature")
# mono <-"Monocyte"
# gran <-c("Eosinophil.Cells","Mast.Cells", "Neutrophil.Cells")
# macro<-c("M0.Macrophage", "M1.Macrophage","M2.Macrophage")

compactPopGroupsList <- list(B_cells=c("B_Cells_Naive","B_Cells_Memory","Plasma_Cells"),
                             CD8cells=c("T_Cells_CD8"),
                             CD4cells=c("T_Cells_CD4_Naive","T_Cells_CD4_Memory_Resting","T_Cells_CD4_Memory_Activated"),
                             Tregs = c("T_Cells_Regulatory_Tregs"),
                             NKcells=c("NK_Cells_Resting","NK_Cells_Activated"),
                             # DCcells=c("Dendritic_Cells_Resting","Dendritic_Cells_Activated"),
                             # Monocytes=c("Monocytes"),
                             Granulocytes=c("Eosinophils_41","Mast_Cells_Resting","Mast_Cells_Activated","Neutrophils_48")
                             # Macrophages=c("Macrophages_M0", "Macrophages_M1","Macrophages_M2")
                             )

compactNames<- names(compactPopGroupsList)



## ABSOLUTE

# Subset to compact sub-populations
resCiber_abs.c <- mitcrDf2.sub[,c(compactPopulations)] %>% as.data.frame()#"bcr_patient_barcode",
# resCiber_abs.c <- resCiber_abs.c %>% column_to_rownames(var="bcr_patient_barcode")

absCompact<- sapply(1:length(compactNames), function(x) sumColumns(resCiber_abs.c,compactNames[x],compactPopGroupsList[[x]]),simplify = FALSE)

absCompact <- bind_cols(absCompact)

mitcrDf2.sub <- cbind(mitcrDf2.sub, absCompact[-2])

colnames(mitcrDf2.sub)[31:64] <- paste0(colnames(mitcrDf2.sub)[31:64],".CIBER")



colnames(danaherImmInfDf.sub.f.agg)[4:17] <- str_remove(colnames(danaherImmInfDf.sub.f.agg)[4:17]," ")
colnames(danaherImmInfDf.sub.f.agg)[4:17] <-  paste0(colnames(danaherImmInfDf.sub.f.agg)[4:17],".DAN")
```


## Merge

```{r,warning=F, message=F, eval=TRUE}
# # Make dataframe with bcr patient barcode project and tcgha.dataset to deal with missing info on those.
# dfToMerge <- list(biospecimen.tpnt.rna.merged.cl.f.sub.f %>% as.data.frame(),
#                participDf.sub.f  %>% as.data.frame(),
#                clinDf.sub  %>% as.data.frame(),
#                tmbDf.sub  %>% as.data.frame(),
#                danaherImmInfDf.sub.f.agg  %>% as.data.frame(),
#                tumorPurityEstimate.sub.f.agg  %>% as.data.frame(),
#                 gene_signatures_rna.sub.f.agg  %>% as.data.frame(),
#                 gene_TISsignature_rna_TP.sub.f  %>% as.data.frame(),
#                 mitcrDf2.sub %>% as.data.frame())
# 
# colsToMerge <- c(colnames(participDf.sub.f)[4:ncol(participDf.sub.f)],
#                  colnames(clinDf.sub)[4:ncol(clinDf.sub)],
#                  colnames(tmbDf.sub)[4:ncol(tmbDf.sub)],
#                  colnames(danaherImmInfDf.sub.f.agg)[4:ncol(danaherImmInfDf.sub.f.agg)],
#                  colnames(tumorPurityEstimate.sub.f.agg)[4:ncol(tumorPurityEstimate.sub.f.agg)],
#                  colnames(gene_signatures_rna.sub.f.agg)[4:ncol(gene_signatures_rna.sub.f.agg)],
#                  colnames(gene_TISsignature_rna_TP.sub.f)[4:ncol(gene_TISsignature_rna_TP.sub.f)],
#                  colnames(mitcrDf2.sub)[4:ncol(mitcrDf2.sub)]
#                  )
# 
# tcga.tp.Filt.final <-rbindlist(dfToMerge, fill=TRUE)[, c("disease_name",
#                  "age_at_initial_pathologic_diagnosis", "gender","race", 
#                  "ajcc_pathologic_tumor_stage","clinical_stage","histological_type", 
#                  "histological_grade","initial_pathologic_dx_year", "menopause_status",
#                   "birth_days_to", "vital_status","tumor_status", 
#                  "last_contact_days_to","death_days_to","cause_of_death",
#                 "new_tumor_event_type","new_tumor_event_site","new_tumor_event_site_other", 
#                 "new_tumor_event_dx_days_to","treatment_outcome_first_course", "margin_status", 
#                 "residual_tumor","OS", "OS.time", 
#                 "DSS", "DSS.time","DFI", 
#                 "DFI.time","PFI","PFI.time",
#                 "Redaction",
#                 "Silent_per_Mb","Non_silent_per_Mb", "logTMB","B-cells.DAN","DC.DAN","Macrophages.DAN","ExhaustedCD8.DAN","T-cells.DAN","CD8T cells.DAN","Neutrophils.DAN","Cytotoxiccells.DAN","Treg.DAN","NKCD56dim cells.DAN", "Mastcells.DAN","NKcells.DAN","CD45.DAN",            "Th1cells.DAN","StromalScore",  "ImmuneScore","ESTIMATEScore", "TumorPurity",
#                  "IFNg_expanded_sigscore", "IFNg_small_sigscore", "TuTACK_sigscore", "PD_L1",
#                  "TIS_sigscore",
#                  "TCGA_Subtype","Immune_Subtype","Leukocyte_Fraction",
#   "Stromal_Fraction","Intratumor_Heterogeneity","TIL_Regional_Fraction",
#   "Proliferation" , "Wound_Healing", "Macrophage_Regulation",
#   "Lymphocyte_Infiltration_Signature_Score", "IFN-gamma_Response", "TGF-beta_Response",
#   "SNV_Neoantigens","Indel_Neoantigens","Silent_Mutation_Rate",
#   "Nonsilent_Mutation_Rate","Number_of_Segments", "Fraction_Altered",
#   "Aneuploidy_Score","Homologous_Recombination_Defects","BCR_Evenness",
#   "BCR_Shannon", "BCR_Richness","TCR_Shannon",
#   "TCR_Richness","TCR_Evenness","CTA_Score",
#   "Th1_Cells.CIBER" ,"Th2_Cells.CIBER","Th17_Cells.CIBER",
#   "B_Cells_Memory.CIBER" ,"B_Cells_Naive.CIBER","Dendritic_Cells_Activated.CIBER",
#   "Dendritic_Cells_Resting.CIBER","Eosinophils_41.CIBER","Macrophages_M0.CIBER",
#   "Macrophages_M1.CIBER","Macrophages_M2.CIBER","Mast_Cells_Activated.CIBER",
#   "Mast_Cells_Resting.CIBER","Monocytes.CIBER","Neutrophils_48.CIBER",
#   "NK_Cells_Activated.CIBER","NK_Cells_Resting.CIBER","Plasma_Cells.CIBER",
#   "T_Cells_CD4_Memory_Activated.CIBER","T_Cells_CD4_Memory_Resting.CIBER","T_Cells_CD4_Naive.CIBER",
#   "T_Cells_CD8.CIBER","T_Cells_Follicular_Helper.CIBER" ,"T_Cells_gamma_delta.CIBER", 
#   "T_Cells_Regulatory_Tregs.CIBER" ,"Lymphocytes.CIBER" ,"Mast_Cells.CIBER",
#   "Dendritic_Cells.CIBER","Macrophages.CIBER","B_cells.CIBER",
#   "CD4cells.CIBER" ,"Tregs.CIBER","NKcells.CIBER","Granulocytes.CIBER"
#                  ) :=lapply(.SD, na.omit) , bcr_patient_barcode, .SDcols=colsToMerge][] %>% distinct()
# 

dfToMerge <- list(biospecimen.tpnt.rna.merged.cl.f.sub.f %>% as.data.frame(),
               participDf.sub.f  %>% as.data.frame(),
               clinDf.sub  %>% as.data.frame(),
               tmbDf.sub  %>% as.data.frame(),
               danaherImmInfDf.sub.f.agg  %>% as.data.frame(),
               tumorPurityEstimate.sub.f.agg  %>% as.data.frame(),
                gene_signatures_rna.sub.f.agg  %>% as.data.frame(),
                gene_TISsignature_rna_TP.sub.f  %>% as.data.frame(),
                mitcrDf2.sub %>% as.data.frame())

colsToMerge <- c(colnames(participDf.sub.f)[4:ncol(participDf.sub.f)],
                 colnames(clinDf.sub)[4:ncol(clinDf.sub)],
                 colnames(tmbDf.sub)[4:ncol(tmbDf.sub)],
                 colnames(danaherImmInfDf.sub.f.agg)[4:ncol(danaherImmInfDf.sub.f.agg)],
                 colnames(tumorPurityEstimate.sub.f.agg)[4:ncol(tumorPurityEstimate.sub.f.agg)],
                 colnames(gene_signatures_rna.sub.f.agg)[4:ncol(gene_signatures_rna.sub.f.agg)],
                 colnames(gene_TISsignature_rna_TP.sub.f)[4:ncol(gene_TISsignature_rna_TP.sub.f)],
                 colnames(mitcrDf2.sub)[4:ncol(mitcrDf2.sub)]
                 )

tcga.tp.Filt.final <-rbindlist(dfToMerge, fill=TRUE)[, c("disease_name",
                 "age_at_initial_pathologic_diagnosis", "gender","race", 
                 "ajcc_pathologic_tumor_stage","clinical_stage","histological_type", 
                 "histological_grade","initial_pathologic_dx_year", "menopause_status",
                  "birth_days_to", "vital_status","tumor_status", 
                 "last_contact_days_to","death_days_to","cause_of_death",
                "new_tumor_event_type","new_tumor_event_site","new_tumor_event_site_other", 
                "new_tumor_event_dx_days_to","treatment_outcome_first_course", "margin_status", 
                "residual_tumor","OS", "OS.time", 
                "DSS", "DSS.time","DFI", 
                "DFI.time","PFI","PFI.time",
                "Redaction",
                "Silent_per_Mb","Non_silent_per_Mb", "logTMB","B-cells.DAN","DC.DAN","Macrophages.DAN","ExhaustedCD8.DAN","T-cells.DAN","CD8T cells.DAN","Neutrophils.DAN","Cytotoxiccells.DAN","Treg.DAN","NKCD56dim cells.DAN", "Mastcells.DAN","NKcells.DAN","CD45.DAN",            "Th1cells.DAN","StromalScore",  "ImmuneScore","ESTIMATEScore", "TumorPurity",
                 "IFNg_expanded_sigscore", "IFNg_small_sigscore", "TuTACK_sigscore", "PD_L1","TLS_score",
                 "TIS_sigscore",
                 "TCGA_Subtype","Immune_Subtype","Leukocyte_Fraction",
  "Stromal_Fraction","Intratumor_Heterogeneity","TIL_Regional_Fraction",
  "Proliferation" , "Wound_Healing", "Macrophage_Regulation",
  "Lymphocyte_Infiltration_Signature_Score", "IFN-gamma_Response", "TGF-beta_Response",
  "SNV_Neoantigens","Indel_Neoantigens","Silent_Mutation_Rate",
  "Nonsilent_Mutation_Rate","Number_of_Segments", "Fraction_Altered",
  "Aneuploidy_Score","Homologous_Recombination_Defects","BCR_Evenness",
  "BCR_Shannon", "BCR_Richness","TCR_Shannon",
  "TCR_Richness","TCR_Evenness","CTA_Score",
  "Th1_Cells.CIBER" ,"Th2_Cells.CIBER","Th17_Cells.CIBER",
  "B_Cells_Memory.CIBER" ,"B_Cells_Naive.CIBER","Dendritic_Cells_Activated.CIBER",
  "Dendritic_Cells_Resting.CIBER","Eosinophils_41.CIBER","Macrophages_M0.CIBER",
  "Macrophages_M1.CIBER","Macrophages_M2.CIBER","Mast_Cells_Activated.CIBER",
  "Mast_Cells_Resting.CIBER","Monocytes.CIBER","Neutrophils_48.CIBER",
  "NK_Cells_Activated.CIBER","NK_Cells_Resting.CIBER","Plasma_Cells.CIBER",
  "T_Cells_CD4_Memory_Activated.CIBER","T_Cells_CD4_Memory_Resting.CIBER","T_Cells_CD4_Naive.CIBER",
  "T_Cells_CD8.CIBER","T_Cells_Follicular_Helper.CIBER" ,"T_Cells_gamma_delta.CIBER", 
  "T_Cells_Regulatory_Tregs.CIBER" ,"Lymphocytes.CIBER" ,"Mast_Cells.CIBER",
  "Dendritic_Cells.CIBER","Macrophages.CIBER","B_cells.CIBER",
  "CD4cells.CIBER" ,"Tregs.CIBER","NKcells.CIBER","Granulocytes.CIBER"
                 ) :=lapply(.SD, na.omit) , bcr_patient_barcode, .SDcols=colsToMerge][] %>% distinct()

table(tcga.tp.Filt.final$project)
median(tcga.tp.Filt.final$TIS_sigscore, na.rm = T)
```

We want to keep the following data:

1. patient ID
2. TCGA project with molecular subtype
3. TCGA project (to see if for some missing)
4. TMB 
5. GEP, TuTack scores & PD-L1 normalized expression
6. MiTCR clones
7. Immune pancancer data: Richness, Evenness, Shannon Entropy and infiltration data
8. Survival data


# Missing values

Visualize variables and missing data across variables.

```{r,warning=F, message=F,fig.show = 'hold', eval = TRUE, fig.width=15, fig.height=12}

visDF <- tcga.tp.Filt.final %>% distinct()

visDF <- visDF %>% distinct()

missmap(visDF, col=c("blue", "Indian red"), legend=FALSE)

```


# Save data

```{r,warning=F, message=F, eval=TRUE}
# save(tcga.tp.nonFilt.final,file=paste0(tcgaClinRDataPath, "tcga.tp.nonFilt.final.RData"))
save(tcga.tp.Filt.final,file=paste0(tcgaIntermediateData, "tcga.tp.Filt.final.RData"))
# save(tcga.tp.barcodes,file=paste0(tcgaClinRDataPath, "tcga.tp.barcodes.Filtered.RData"))

```


# REMOVE ICI DATA


```{r,warning=F, message=F}
mycancertypes <- levels(as.factor(tcga.tp.Filt.final$project))

## Exclude THYM and DLBC
mycancertypes.reduced <- mycancertypes[-which(mycancertypes %in% c("COAD_MSI_low","BRCA_Her2","BRCA_LuminalA", "BRCA_LuminalB","BRCA_Normal","BRCA_TN","COAD_MSI","COAD_MSS"))]

mycancertypes.multi <- mycancertypes.reduced[mycancertypes.reduced %in% c("SKCM","KIRC", "STAD", "BLCA")]

## Subset tables
tcga.tp.Filt.final.sub <- tcga.tp.Filt.final %>% dplyr::filter(project %in% mycancertypes.reduced)
```

## Download ICI info

Download clinical data and treatment info from TCGAbiolinks for SKCM (melanoma), KIRC and KIRP (RCC), STAD (gastric), BLCA (bladder)


```{r,warning=F, message=F, eval=TRUE, echo=FALSE, results=FALSE}

# # Download clinical data for TCGA
# tcga.drug <- sapply(mycancertypes.multi , function(x) get_clinicalTCGA(x, tcgaDF=tcga.tp.Filt.final,barcodeCol="bcr_patient_barcode"), simplify = FALSE )
# 
# **NOTE**: Possible error arising from TCGAbiolinks package. Due to updates, the data types in the columns might change and the code below might require manual curation so that it runs smoothly. This is quite unpredictable
# 
# # Issue with regimen_number, total_dose, prescribed_dose column that is different data type across the cohorts
# tcga.drug <- lapply(tcga.drug, function(df) {
#   df$regimen_number <- as.character(df$regimen_number)
#   df$total_dose <- as.character(df$total_dose)
#   df$prescribed_dose <- as.character(df$prescribed_dose)
#   
#   return(df)
# })
# 
# tcga.drug.df <-bind_rows(tcga.drug, .id = "column_label")
# colnames(tcga.drug.df)[1] <- "project"

## SKCM
skcm <- tcga.tp.Filt.final %>% filter(project=="SKCM") %>% droplevels()
query <- GDCquery(project = "TCGA-SKCM", 
              data.category = "Clinical", 
              file.type = "xml", 
              barcode = skcm$bcr_patient_barcode)
GDCdownload(query)
clinical.skcm <- GDCprepare_clinic(query, clinical.info = "drug")

# dim(clinical.skcm)
## KIRC and KIRP
rcc <- tcga.tp.Filt.final %>% filter(project %in% c("KIRC")) %>% droplevels()
query <- GDCquery(project = "TCGA-KIRC", 
              data.category = "Clinical", 
              file.type = "xml", 
              barcode = rcc$bcr_patient_barcode)
GDCdownload(query)
clinical.rcc1 <- GDCprepare_clinic(query, clinical.info = "drug")

# rcc <- tcga.tp.Filt.final %>% filter(project %in% c("KIRP")) %>% droplevels()
# query <- GDCquery(project = "TCGA-KIRP", 
#               data.category = "Clinical", 
#               file.type = "xml", 
#               barcode = rcc$bcr_patient_barcode)
# GDCdownload(query)
# clinical.rcc2 <- GDCprepare_clinic(query, clinical.info = "drug")
# 
# clinical.rcc <- rbind(clinical.rcc1,clinical.rcc2)
clinical.rcc <- rbind(clinical.rcc1)
# dim(clinical.rcc)
## STAD
gastric <- tcga.tp.Filt.final %>% filter(project=="STAD") %>% droplevels()
query <- GDCquery(project = "TCGA-STAD", 
              data.category = "Clinical", 
              file.type = "xml", 
              barcode = gastric$bcr_patient_barcode)
GDCdownload(query)
clinical.gastric <- GDCprepare_clinic(query, clinical.info = "drug")
# dim(clinical.gastric)
## BLCA
bladder <- tcga.tp.Filt.final %>% filter(project=="BLCA") %>% droplevels()
query <- GDCquery(project = "TCGA-BLCA", 
              data.category = "Clinical", 
              file.type = "xml", 
              barcode = bladder$bcr_patient_barcode)
GDCdownload(query)
clinical.bladder <- GDCprepare_clinic(query, clinical.info = "drug")
# dim(clinical.bladder)
tcga.drug <- rbind(clinical.skcm,clinical.gastric, clinical.rcc, clinical.bladder)

save(tcga.drug, file=paste0(tcgaIntermediateData,"tcga.drug",Rdata.suffix))
```

## FIND NON ICI SAMPLES

We want to exclude immunotherapies but not IFN-gamma, IFN-alpha, beta, IL-2

```{r,warning=F, message=F}

# SKCM, STAD, BRCA_Her2 DLBC COAD_MSI
## Now subset tcga data to those patients that drug information
tcga.drug <-loadRData(paste0(tcgaIntermediateData,"tcga.drug",Rdata.suffix))
levels(tcga.drug$therapy_types)
unique(as.character(tcga.drug$drug_name))

table(tcga.drug$project)

# DRUGS TO EXCLUDE
tcga.drug.ici <- tcga.drug %>% dplyr::filter(therapy_types %in% c("Immunotherapy")) %>% droplevels()
unique(as.character(tcga.drug.ici$drug_name))
ici_to_exclude <- c("ipilimumab" , "Pembrolizumab", "Ipilimumab")

tcga.drug.ici.To.Exclude <- tcga.drug %>% dplyr::filter(therapy_types %in% c("Immunotherapy")) %>%
                                          dplyr::filter(drug_name %in% ici_to_exclude) %>% droplevels() %>%
                                          pull(bcr_patient_barcode)

length(tcga.drug.ici.To.Exclude)


tcga.drug.ici.To.Exclude2 <- tcga.drug %>% dplyr::filter(drug_name %in% ici_to_exclude) %>% droplevels() %>%
                                          pull(bcr_patient_barcode)

length(tcga.drug.ici.To.Exclude2)

tcga.drug.ici.Only <- tcga.drug %>% dplyr::filter(bcr_patient_barcode %notin% tcga.drug.ici.To.Exclude2) %>% droplevels()

table(tcga.drug.ici.Only$project)

```

## Save data

```{r,warning=F, message=F}
# save(tcga.drug.ici.Only,file=paste0(tcgaIntermediateData, "tcga.tp.Filt.final_ICI4Tissues.RData"))

save(tcga.drug.ici.To.Exclude2,file=paste0(tcgaIntermediateData, "tcga_patientsNOTici_toEXCLUDE_4Tissues.RData"))
```


# Session

```{r,warning=F, message=F,fig.show = 'hold', eval = TRUE, fig.width=15, fig.height=14}
# Cleanup
unlink(paste0("GDCdata"),recursive=TRUE)
unlink(paste0("MANIFEST.txt"),recursive=TRUE)
unlink(paste0('*VennDiagram*'))
# Session
session_info <- sessionInfo()
writeLines(capture.output(session_info), paste0(sessionInfoPath,"A_06_tcga_clinical_biomarker_metadata.txt"))

sessionInfo()

```